---
title: "hw1"
format: pdf
editor: visual
---

# Homework 1

## Problem 1

#### (a) Import the data and change column names.

```{r}
abal<-read.csv("abalone.data",header=FALSE)
colnames(abal)<-c("sex","len","diam","ht","whole_wt","shu_wt","vis_wt","shel_wt","rings")
abal
```

#### (b) Numbers of observations belong to each sex

```{r}
#| echo: false
table(abal$sex)
```

#### (c) correlation

1.From the result below, shell weight has the highest correlation with rings.

```{r}
weight<-abal[,c("whole_wt","shu_wt","vis_wt","shel_wt")]
corr<-cor(weight, abal$rings)
corr
```

2.As shown below, the infant group has the highest correlation between shell weight and rings.

```{r}
ab_sex<-c("M","F","I")
corr_groupby_sex<-sapply(ab_sex,function(x) cor(abal$shel_wt[abal$sex==x],abal$rings[abal$sex==x]))
print(corr_groupby_sex)
```

3.Four weights of abalone with the highest rings.

```{r}
max_ring<-max(abal$rings)
wt_with_max_ring<-abal[abal$rings==max_ring,c("whole_wt","shu_wt","vis_wt","shel_wt")]
wt_with_max_ring
```

4.The percentage of abalones whose viscera weight larger than their shell weight is apprximately 6.51%.

```{r}
target<-abal[abal$vis_wt>abal$shel_wt,]
percent<-nrow(target)/nrow(abal)
percent
```

#### (d) correlation table group by sex

```{r}
library(dplyr)
cor_table <- abal %>%group_by(sex) %>%summarize(
              whole_wt = cor(whole_wt, rings),
    shu_wt = cor(shu_wt, rings),
    vis_wt = cor(vis_wt, rings),
    shel_wt = cor(shel_wt, rings)
  )

print(cor_table)
```

#### (e) t test

1.t test between F&I: p value\<0.05, indicating there are significant difference of rings between female and infant abalones under significance level=0.05

```{r}
FI <- t.test(rings ~ sex, data = abal, subset = sex %in% c("F", "I"))
FI
```

2.t test between F&M p value\<0.05, indicating there are significant difference of rings between female and male abalones under significance level=0.05

```{r}
FM <- t.test(rings ~ sex, data = abal, subset = sex %in% c("F", "M"))
FM
```

3.t test between I&M p value\<0.05, indicating there are significant difference of rings between infant and male abalones under significance level=0.05

```{r}
IM <- t.test(rings ~ sex, data = abal, subset = sex %in% c("I", "M"))
IM
```

## Problem 2

#### (a) Import dataset.

```{r}
dt<-read.csv("food_expenditure.csv")
dt
```

#### (b) Revise Column Names.

```{r}
colnames(dt)<-c("id","age","fam_meb","state","currency","foodexp","foodexp_gro",
                "foodexp_dinout","foodexp_mis","dinout_freq","alco_include",
                "ass_prog")

```

#### (c) Restrict currency to USD.

After resricting currency, the total number of rows reduced to 230.

```{r}
print(nrow(dt)) #number of observations before restricting the data.
data<-dt[dt$currency=="USD",]
print(nrow(data))#number of observations after restricting the data.
```

#### (d) Age

```{r}
sum(is.na(data$age))
```

restricting age \>=18 and eliminating missing values

```{r}
data1<-data[!is.na(data$age) & data$age>=18,]
```

#### (e) State

```{r}
sum(is.na(data1$state))
```

```{r}
nrow(data1[data1$state=='XX',])
```

Eliminate the rows with state='XX' or missing state values

```{r}
data2<-data1[data1$state!='XX',]
```

#### (f) Food expenditure

Exlude missing values and the values smaller than or equal to 0.

```{r}
sum(is.na(data2$foodexp))
```

```{r}
sum(is.na(data2$foodexp_dinout))
```

```{r}
sum(is.na(data2$foodexp_gro))
```

```{r}
sum(is.na(data2$foodexp_mis))
```

```{r}
data3<-data2[!is.na(data2$foodexp_dinout) & !is.na(data2$foodexp_gro) &     
               !is.na(data2$foodexp_mis) & data2$foodexp_dinout>0 & data2$foodexp_gro>0
              & data2$foodexp_mis>0 & data2$foodexp>0,]
```

#### (g) Dine-out frequency

eliminate empty values and make sure dinout_freq\>=0

```{r}
data4<-data3[!is.na(data3$dinout_freq) & data3$dinout_freq>=0,]
```

#### (h) After filtering, there are 128 valid observations.

```{r}
nrow(data4)
```

```{r}
data4
```

## Problem 3

#### (a) Build function return collatz number

```{r}
# Function to give the next collatz
nextCollatz<-function(x){
  #inputs: x- an integer
  #outputs: its next collatz number
  if (x<1) {return(FALSE)}
  if (x==1){return(1)}
  else{
    if (x%%2==0){return (x/2)}
    else{return(3*x+1)}
    }
  
  
}
```

```{r}
nextCollatz(5)
```

```{r}
nextCollatz(16)
```

```{r}
nextCollatz(-1)
```

#### (b) Build function return collatz sequence

```{r}
collatzSequence<-function(x){
  #inputs: x- an integer
  #outputs: the sequence of collatz number starting at x
  if (x==1){return(1)}
  if(x<1){return(FALSE)}
  if(x>1){
    res<-c(x)
    while(x>1){
      res<-append(res,nextCollatz(x))
      x<-nextCollatz(x)
    }
  return(res)}
}
```

```{r}
collatzSequence(5)
```

```{r}
collatzSequence(0)
```

```{r}
collatzSequence(100)
```

#### (c) Finding the longest and shortest sequences

```{r}
df<-data.frame(start=integer(),len=integer())
for (i in 100:500){
  new_row<-data.frame(start=i,len=length(collatzSequence(i)))
  df <- rbind(df, new_row)
}
```

Longest sequence

```{r}
max_len<-max(df['len'])
max_row<-df[df$len==max_len,]
max_row
```

Shortest sequence

```{r}
min_len<-min(df['len'])
min_row<-df[df$len==min_len,]
min_row
```
