---
title: "hw3"
format: pdf
editor: visual
---

## Problem 1

#### a.

```{r}
library(haven)
aux<-read_xpt("AUX_I.xpt")
demo<-read_xpt("DEMO_I.xpt")
data<-merge(aux, demo, by = "SEQN") 
dim(data)
```

#### b.

```{r}
#RIAGENDR--GENDER
#DMDCITZN--Citizenship status
#DMDHHSZE--Number of children 5 years or younger 
#INDHHIN2--Annual household income 
vars<-c("RIAGENDR", "DMDCITZN", "DMDHHSZE", "INDHHIN2")
data$gender<-factor(data$RIAGENDR,levels = c(1, 2),labels = c("Male", "Female"))

data$citizen_sta<-factor(data$DMDCITZN,levels=c(1,2,7,9),
             labels=c("citizen","noncitizen",NA,NA))
data$child_under5<-ordered(data$DMDHHSZE,levels=c(0,1,2,3),labels=c('0','1','2','3 or more'))
data$house_income<-ordered(data$INDHHIN2,levels=append(1:10,c(12,12,14,15,77,99)),
                      labels=c('$ 0 to $ 4,999','$ 5,000 to $ 9,999',
                               '$10,000 to $14,999', '$15,000 to $19,999',
                               '$20,000 to $24,999','$25,000 to $34,999',
                               '$35,000 to $44,999','	$45,000 to $54,999',
                               '$55,000 to $64,999','$65,000 to $74,999',
                               NA,NA,
                               '$75,000 to $99,999','$100,000 and Over',
                               NA,NA))
```

#### c.

Fit Poisson regression model.

```{r}
model_1R<-glm(AUXTWIDR~gender,data,family = poisson)
model_2R<-glm(AUXTWIDR~gender+citizen_sta+child_under5+
              house_income,data,family=poisson)
model_1L<-glm(AUXTWIDL~gender,data,family = poisson)
model_2L<-glm(AUXTWIDL~gender+citizen_sta+child_under5+
                house_income,data, family=poisson)
```



sample size, pseudo-R square, and AIC values

```{r}
library(knitr)
library(broom)
library(tidyverse)
library(dplyr)
model_stats<-function(model,name) {
  pseudo_r2<-1-(model$deviance/model$null.deviance)
  data.frame(
    Model=name,
    N=nobs(model),
    Pseudo_R2 = round(pseudo_r2,3),
    AIC = round(AIC(model),1)
  )
}
md_stats<-bind_rows(
  model_stats(model_1R, "1R"),
  model_stats(model_2R, "2R"),
  model_stats(model_1L, "1L"),
  model_stats(model_2L, "2L")
)
kable(md_stats)
```

IRR

```{r}
get_irr<-function(model,name) {
  coef_summary <- summary(model)$coefficients
  irr<-exp(coef_summary[,"Estimate"])
  p_values<-coef_summary[,"Pr(>|z|)"]
  results <- data.frame(
    model=name,
    IRR = round(irr,3),
    p_value = round(p_values,3))
  return(results)
}
irr_table <- bind_rows(
  get_irr(model_1R,"1R"),
  get_irr(model_2R,"2R"),
  get_irr(model_1L,"1L"),
  get_irr(model_2L,"2L")
)
kable(irr_table)
```

#### d.
Test for IRR:
The p-value for genderFemale in model 2L is $8.25*10^{-4}$, which means gender is a significant at α=0.01,and irr greater than 1, means female have more  Tympanometric width measure than male.

```{r}
res<-summary(model_2L)$coefficients
irr<-exp(res[,"Estimate"])
res
```
```{r}
irr
```
Test for predicted value:
since p value<0.01, so reject null hypothesis under α=0.01, predicted value of Tympanometric width measure of the left ear differs between men and women.

```{r}
data$pred<-predict(model_2L, newdata=data)
t.test(pred~gender,data=data)
```
## Problem 2

Loading data:
```{r}
library(microbenchmark)
library(DBI)
sakila <- dbConnect(RSQLite::SQLite(), "sakila_master.db")
dbListTables(sakila)
```
#### a.
###### 1)SQL+R:
SQL:
```{r}
customers<-dbGetQuery(sakila,"
        SELECT customer_id, store_id, active
          FROM customer;         
        ")
```
R:
```{r}
library(dplyr)
actcust<- customers %>%
  group_by(store_id) %>%
  summarize(
    n_customers = n(),
    percent_active =sum(as.integer(active))/n_customers*100
  ) 
print(actcust)
```
###### 2) Single SQL query
```{r}
dbGetQuery(sakila,"
        SELECT store_id,COUNT(*) AS n_customers,
               100.0 * SUM( active ) / COUNT(*) AS percent_active
            FROM customer 
            GROUP BY store_id;
")
```
###### 3) time comparison
It seems that using only SQL query is faster.
```{r}
bench <- microbenchmark(
  SQL_R= {
    customers <- dbGetQuery(sakila, "SELECT customer_id, store_id, active FROM customer")
    customers %>%
      group_by(store_id) %>%
      summarize(n_customers = n(),
                n_active = sum(as.integer(active)),
                pct_active = n_active / n_customers * 100)
  },
  SQL= {
        dbGetQuery(sakila,"
        SELECT store_id,COUNT(*) AS n_customers,
               100.0 * SUM( active ) / COUNT(*) AS percent_active
            FROM customer 
            GROUP BY store_id;
")
  },
  times = 20
)
print(bench)
```
#### b.
SQL+Rmethod has relatively shorter time.
###### 1)SQL+R
```{r}
staf<-dbGetQuery(sakila, "
  SELECT
    s.staff_id,
    s.first_name,
    s.last_name,
    a.city_id,
    a.address_id,
    a.address,
    c.city,
    c.country_id
  FROM staff AS s
  JOIN address AS a ON s.address_id = a.address_id
  JOIN city AS c ON a.city_id = c.city_id
")

ctr<-dbGetQuery(sakila, "
  SELECT country_id, country
   FROM country
")
staff_info<-staf%>%
  left_join(ctr, by = "country_id") %>%
  select(staff_id, first_name, last_name, country)
staff_info
```
###### 2)single SQL
```{r}
dbGetQuery(sakila, "
  SELECT 
    s.staff_id,
    s.first_name,
    s.last_name,
    a.city_id,
    ct.city,
    cr.country_id
  FROM staff AS s
  JOIN address AS a ON s.address_id = a.address_id
  JOIN city AS ct ON a.city_id = ct.city_id
  JOIN country cr ON ct.country_id = cr.country_id;
")
```
###### 3)
```{r}
bench <- microbenchmark(
  SQL_R= {
staf<-dbGetQuery(sakila, "
  SELECT
    s.staff_id,
    s.first_name,
    s.last_name,
    a.city_id,
    a.address_id,
    a.address,
    c.city,
    c.country_id
  FROM staff AS s
  JOIN address AS a ON s.address_id = a.address_id
  JOIN city AS c ON a.city_id = c.city_id
")

ctr<-dbGetQuery(sakila, "
  SELECT country_id, country
   FROM country
")
staff_info<-staf%>%
left_join(ctr, by = "country_id") %>%
select(staff_id, first_name, last_name, country)
staff_info
  },
  SQL= {
dbGetQuery(sakila, "
  SELECT 
    s.staff_id,
    s.first_name,
    s.last_name,
    a.city_id,
    ct.city,
    cr.country_id
  FROM staff AS s
  JOIN address AS a ON s.address_id = a.address_id
  JOIN city AS ct ON a.city_id = ct.city_id
  JOIN country cr ON ct.country_id = cr.country_id;
")
  },
  times = 20
)
print(bench)
```

#### c.
###### 1)SQL+R
```{r}
pay<-dbGetQuery(sakila, "SELECT payment_id, rental_id, amount FROM payment")
rent<-dbGetQuery(sakila, "SELECT rental_id, inventory_id FROM rental")
inv <-dbGetQuery(sakila, "SELECT inventory_id, film_id FROM inventory")
film<-dbGetQuery(sakila, "SELECT film_id, title FROM film")
library(dplyr)
revenue<-pay%>%
  left_join(rent,by="rental_id")%>%
  left_join(inv,by="inventory_id") %>%
  left_join(film,by="film_id") %>%
  group_by(film_id,title) %>%
  summarize(total_revenue=sum(amount,na.rm=TRUE))
max_revenue <- max(revenue$total_revenue, na.rm = TRUE)
top<-revenue %>%
  filter(total_revenue==max_revenue)
print(top)


```


###### 2)SQL
`
```{r}
dbGetQuery(sakila,'
       SELECT f.title, SUM(p.amount) AS total_revenue
         FROM film f
          JOIN inventory i ON f.film_id = i.film_id
          JOIN rental r ON i.inventory_id = r.inventory_id
          JOIN payment p ON r.rental_id = p.rental_id
        GROUP BY f.film_id, f.title
        HAVING total_revenue = (SELECT MAX(total) FROM (SELECT SUM(p.amount) AS total 
        FROM payment p 
        JOIN rental r ON p.rental_id = r.rental_id 
        JOIN inventory i ON r.inventory_id = i.inventory_id 
        GROUP BY i.film_id) AS sub);')
```

###### 3)
It seems that SQL+R method is more efficient.

```{r}
bench <- microbenchmark(
  SQL_R= {
    pay<-dbGetQuery(sakila, "SELECT payment_id, rental_id, amount FROM payment")
   rent<-dbGetQuery(sakila, "SELECT rental_id, inventory_id FROM rental")
   inv <-dbGetQuery(sakila, "SELECT inventory_id, film_id FROM inventory")
   film<-dbGetQuery(sakila, "SELECT film_id, title FROM film")
   library(dplyr)
   revenue<-pay%>%
  left_join(rent,by="rental_id")%>%
  left_join(inv,by="inventory_id") %>%
  left_join(film,by="film_id") %>%
  group_by(film_id,title) %>%
  summarize(total_revenue=sum(amount,na.rm=TRUE))
  max_revenue <- max(revenue$total_revenue, na.rm = TRUE)
  top<-revenue %>%
    filter(total_revenue==max_revenue)
  },
  SQL= {
        dbGetQuery(sakila,'
       SELECT f.title, SUM(p.amount) AS total_revenue
         FROM film f
          JOIN inventory i ON f.film_id = i.film_id
          JOIN rental r ON i.inventory_id = r.inventory_id
          JOIN payment p ON r.rental_id = p.rental_id
        GROUP BY f.film_id, f.title
        HAVING total_revenue = (SELECT MAX(total) FROM (SELECT SUM(p.amount) AS total 
        FROM payment p 
        JOIN rental r ON p.rental_id = r.rental_id 
        JOIN inventory i ON r.inventory_id = i.inventory_id 
        GROUP BY i.film_id) AS sub);')
  },
  times = 20
)
print(bench)
```



## Problem 3
#### a.
```{r}
data<-read.csv("au_500.csv")
data

```
```{r}
com_web<-grepl("\\.com$",data$web,ignore.case=TRUE)
percent_com<-mean(com_web, na.rm = TRUE)
percent_com
```
#### b.
```{r}
# 提取电子邮件域名
email_domains <- sub(".*@", "", data$email)

# 找出频率最高的域名
domain_freq <- sort(table(email_domains), decreasing = TRUE)
cat("最常见的电子邮件域名是：", names(domain_freq)[1], "（共", domain_freq[1], "个）\n")

```
#### c.
```{r}
# 移除空格和逗号后检查是否有非字母字符
name_cleaned <- gsub("[ ,]", "", data$company_name)

# 含有非字母字符的公司名
non_alpha <- grepl("[^A-Za-z]", name_cleaned)
proportion_non_alpha <- mean(non_alpha) * 100
cat("包含非字母字符（不含逗号、空格）的公司名称比例：", round(proportion_non_alpha, 2), "%\n")

# 如果也排除 & 字符
name_cleaned2 <- gsub("[ ,&]", "", data$company_name)
non_alpha2 <- grepl("[^A-Za-z]", name_cleaned2)
proportion_non_alpha2 <- mean(non_alpha2) * 100
cat("包含非字母字符（不含逗号、空格、&）的公司名称比例：", round(proportion_non_alpha2, 2), "%\n")

```
#### d.
```{r}
# 创建格式化函数
fmt_phone <- function(ph) {
  # 去掉非数字字符
  dig <- str_replace_all(ph, "[^0-9]", "")
  # 若长度不是 10，则返回 NA 或原样
  if (nchar(dig) != 10) {
    return(NA)
  }
  # 切分：前 4，次 3，后 3
  p1 <- substr(dig, 1, 4)
  p2 <- substr(dig, 5, 7)
  p3 <- substr(dig, 8, 10)
  return(paste(p1, p2, p3, sep = "-"))
}

data$phone1_fmt <- vapply(data$phone1, fmt_phone, FUN.VALUE = character(1))
data$phone2_fmt <- vapply(data$phone2, fmt_phone, FUN.VALUE = character(1))

# 打印前 10
head(data$phone1_fmt, 10)
head(data$phone2_fmt, 10)


```


