---
title: "Homework 4"
format: pdf
editor: visual
---

## Problem 1

```{r}
library(tidyverse)
library(nzelect)
data(nzge)
```

#### (a)

```{r}
#| echo: false
resulta<-nzge %>%
  group_by(election_year,voting_type) %>%
  summarise(count=sum(votes)) %>%
  arrange(count) %>%
  ungroup()
print(resulta)
```

#### (b)

```{r}
resultb<-nzge %>%
  filter(election_year=='2014') %>%
  group_by(party) %>%
  summarise(votesum=sum(votes)) %>%
  mutate(percent=votesum/sum(votesum)*100) %>%
  select(!votesum) %>%
  arrange(desc(percent))
print(resultb)
```

#### (c)

```{r}
resultc<-nzge %>%
  group_by(election_year,voting_type,party) %>%
  summarise(total=sum(votes)) %>%
  group_by(election_year,voting_type) %>%
  slice_max(order_by=desc(total),n=1,with_ties = FALSE) %>%
  ungroup()%>%
    select(!total) %>%
  pivot_wider(
    names_from = voting_type,
    values_from = party,
    names_prefix = "Winner_"
  ) %>%
  arrange(election_year)
print(resultc)
```

## Problem 2

Load the data

```{r}
tennis<-readr::read_delim("https://raw.githubusercontent.com/JeffSackmann/tennis_atp/refs/heads/master/atp_matches_2019.csv")
```

#### (a)

There are 125 tournments took place in 2019.

```{r}
tour_2019<- tennis %>%
  filter(tourney_date>=20190000) %>%
  distinct(tourney_id, tourney_name, tourney_date) %>%
  count()
print(as.numeric(tour_2019))
```

#### (b)

Yes. There are 25 players won more than 1 game.

```{r}
tourn_winner<-tennis %>%
  group_by(tourney_id,tourney_name) %>%
  summarise(winner=first(winner_name)) %>%
  ungroup() %>%
  count(winner,name="tourn_win") %>%
  filter(tourn_win>1) %>%
  arrange(desc(tourn_win))
tourn_winner %>% nrow
```

The most winning players won 7 tournments.

```{r}
most_win<-tourn_winner %>%
  first
most_win
```

#### (c)

$H_0:\mu_{w-ace}=\mu_{l-ace}$ $H_1:\mu_{w-ace}>\mu_{l-ace}$ Since p value\<0.05, we reject H0 under significance level 0.05,there is any evidence that winners have more aces than losers.

```{r}
aces<-tennis %>%
  select(winner_name,loser_name,w_ace,l_ace) %>%
  drop_na(w_ace,l_ace) %>%
  summarize(
    mean_diff=mean(w_ace)-mean(l_ace),
    test_res=list(t.test(w_ace,l_ace,paired=TRUE,alternative = "greater"))
  )
p_value<-aces$test_res[[1]]$p.value
print(p_value)
```

#### (d)
Rafael Nadal has the highest win rate among the players who palyed at least 5 matches.
```{r}
highest_win_rate<-tennis %>%
  select(winner_name,loser_name) %>%
  pivot_longer(cols=c(winner_name,loser_name),
               names_to='result',
               values_to='player') %>%
  mutate(win=ifelse(result=='winner_name',1,0)) %>%
  group_by(player) %>%
  summarise(total=n(),
            win_time=sum(win),
            win_rate=win_time/total) %>%
  filter(total>=5) %>%
  arrange(desc(win_rate)) %>%
  first
print(highest_win_rate)
```

## Problem 3

#### (a)
Daily cases vs. date
```{r}
dt<-read.csv("https://raw.githubusercontent.com/nytimes/covid-19-data/refs/heads/master/rolling-averages/us-states.csv")
dt$date<-as.Date(dt$date)
with(dt,plot(cases~date))
```
Plot monthly cases against month.
```{r}
monthly_dt<-dt %>%
  mutate(month=floor_date(date, "month")) %>%       
  group_by(month) %>%
  summarise(monthly_cases=sum(cases,na.rm = TRUE))
with(monthly_dt,plot(monthly_cases~month))
with(monthly_dt,lines(monthly_cases~month))
```
From the above plots we can conclude that there is 1 major spike and 3 minor spikes in cases.
#### (b)
Identify the states with the highest and lowest overall rates per population
```{r}
state_total_rate<-dt %>%
  group_by(state) %>%
  summarise(total_rate_per_pop=sum(cases_avg_per_100k,na.rm=TRUE))%>%
  arrange(desc(total_rate_per_pop))
high_state<-state_total_rate$state[1]
low_state<-state_total_rate$state[nrow(state_total_rate)]
cat("Highest:",high_state, "\nLowest:",low_state, "\n")
```
 Trajectories over time: Rhode Island fluctuates more obviously than Maine, Rhode Island also has a main spike, which is way higher than Maine.
```{r}
selected_states <- dt %>%
  filter(state %in% c(high_state, low_state))
ggplot(selected_states, aes(x=date,y=cases_avg_per_100k,color=state)) +
  geom_line(size = 1) 
```
#### (c)
The first five states to experience Covid in a substantial way are New York, New Jersey,
Louisiana, Guam and Rhode Island.
```{r}
first_substantial<-dt %>%
  filter(cases_avg_per_100k >= mean(cases_avg_per_100k)) %>% 
  group_by(state) %>%
  summarise(first_date = min(date,na.rm = TRUE)) %>%
  arrange(first_date)
head(first_substantial,5)
```
