---
title: "Homework 5"
format: pdf
editor: visual
---

## Problem 1

#### (a)

###### 1.constructor

```{r}
setClass("waldCI",
         slots = c(lb = 'numeric', 
                   ub = 'numeric', 
                   mean = 'numeric', 
                   sterr = 'numeric', 
                   level = 'numeric'))
WaldCI<-function(lb=NULL,ub=NULL,mean=NULL,sterr=NULL,level){
  if (!is.null(mean) && !is.null(sterr)) {
    z <- .zval(level)
    lb <- mean - z * sterr
    ub <- mean + z * sterr
  } else if (!is.null(lb) && !is.null(ub)) {
    mean <- (lb + ub) / 2
    z <- .zval(level)
    sterr <- (ub - lb) / (2 * z)
  } else {
    stop("must specify either (mean, sterr) or (lb, ub).")
  }
  return(new("waldCI",lb=lb,ub=ub,mean=mean,sterr=sterr,level=level))
}
```

###### 2.validator

```{r}
setValidity("waldCI", function(object){
  if (object@sterr<0 && is.null(object@sterr)) {
    stop("standard error should be positive")
  }
  if (object@ub<=object@lb){
    stop("upper bound of CI should be greater than lower bound")
  }
  if (!is.finite(object@lb) || !is.finite(object@ub)){
    stop("bounds must be finite.")
  }
  if (!is.numeric(object@level) || object@level<=0 || object@level>=1){
    stop("level must be between 0 and 1")}
  return(TRUE)
})
```

###### 3.show method

```{r}
setMethod("show", "waldCI",
  function(object) {
    cat("Wald Confidence Interval (", object@level * 100, "%)\n", sep = "")
    cat("Mean:", object@mean, "\n")
    cat("SE:", object@sterr, "\n")
    cat("CI: (", object@lb, ", ", object@ub, ")\n", sep = "")
    return(invisible(object))
    }
)
```

###### 4.accessors

```{r}
setGeneric("lb", function(object) standardGeneric("lb"))
setGeneric("ub", function(object) standardGeneric("ub"))
setGeneric("mean", function(object) standardGeneric("mean"))
setGeneric("sterr", function(object) standardGeneric("sterr"))
setGeneric("level", function(object) standardGeneric("level"))
setMethod("lb", "waldCI", function(object) object@lb)
setMethod("ub", "waldCI", function(object) object@ub)
setMethod("mean", "waldCI", function(object) object@mean)
setMethod("sterr", "waldCI", function(object) object@sterr)
setMethod("level", "waldCI", function(object) object@level)
```

###### 5.setters

```{r}
setGeneric("lb<-", function(object,value) standardGeneric("lb<-"))
setGeneric("ub<-", function(object,value) standardGeneric("ub<-"))
setGeneric("mean<-", function(object,value) standardGeneric("mean<-"))
setGeneric("sterr<-", function(object,value) standardGeneric("sterr<-"))
setGeneric("level<-", function(object,value) standardGeneric("level<-"))
setMethod("lb<-", "waldCI", function(object,value){
  object@lb<-value
  validObject(object)
  return(object)
} )
setMethod("ub<-", "waldCI", function(object,value){
  object@ub<-value
  validObject(object)
  return(object)
})
setMethod("mean<-", "waldCI", function(object,value){ 
  object@mean<-value
  validObject(object)
  return(object)
})
setMethod("sterr<-", "waldCI", function(object,value){
  object@sterr<-value
  validObject(object)
  return(object)
} )
setMethod("level<-", "waldCI", function(object,value){
  object@level<-value
  validObject(object)
  return(object)
})
```

###### 6.contains

```{r}
setGeneric("contains", function(object, value) standardGeneric("contains"))
setMethod("contains", "waldCI", function(object, value){
  return(value >= object@lb && value <= object@ub)
})
```

###### 7.overlap

```{r}
setGeneric("overlap", function(ci1, ci2) standardGeneric("overlap"))
setMethod("overlap", "waldCI", function(ci1, ci2){
  return(!(ci1@ub<ci2@lb || ci2@ub<ci1@lb))
})
```

###### 8.as.numeric

```{r}
setMethod("as.numeric","waldCI",function(x) c(x@lb,x@ub))
```

###### 9.transformCI

```{r}
setGeneric("transformCI",function(ci,f) standardGeneric("transformCI"))
setMethod("transformCI", "waldCI", function(ci,f) {
  warning("Only monotonic functions make sense.")
  new_lb<-f(ci@lb)
  new_ub<-f(ci@ub)
  WaldCI(level=ci@level,lb=new_lb,ub=new_ub)
})
```

#### (b)

```{r}
ci1<-WaldCI(level=0.95,lb=17.2,ub=24.7)
ci2<-WaldCI(level=0.99,mean=13,sterr=2.5)
ci3<-WaldCI(level=0.75,lb=27.43,ub=39.22)
```

```{r}
ci1
ci2
ci3
as.numeric(ci1)
as.numeric(ci2)
as.numeric(ci3)
lb(ci2)
ub(ci2)
mean(ci1)
sterr(ci3)
level(ci2)
lb(ci2) <- 10.5
mean(ci3) <- 34
level(ci3) <- .8
contains(ci1, 17)
contains(ci3, 44)
overlap(ci1, ci2)
eci1 <- transformCI(ci1, sqrt)
eci1
mean(transformCI(ci2, log))
```

#### (c)

##### 1.negative standard error

```{r}
try(WaldCI(level = 0.95, mean = 10, sterr = -3))
```

##### 2.lb \> ub

```{r}
try(WaldCI(level = 0.95, ub=5,lb=10))
```

##### 3.Infinite bounds

```{r}
try(WaldCI(level = 0.95, ub=Inf,lb=10))
```

##### 4.invalid use of the setters

```{r}
ci<- ci1
try(lb(ci) <- 30)
```

## Problem 3

#### (a)

There are 5 major spikes and 2 minor spikes.

```{r}
library(plotly)
dt<-read.csv("https://raw.githubusercontent.com/nytimes/covid-19-data/refs/heads/master/rolling-averages/us-states.csv")
dt$date<-as.Date(dt$date)
p <- plot_ly(dt, x = ~ date, y = ~ cases, 
             mode = "lines")
```

```{r}
p
```

#### (b)

```{r}
library(tidyverse)
state_total_rate<-dt %>%
  group_by(state) %>%
  summarise(total_rate_per_pop=sum(cases_avg_per_100k,na.rm=TRUE))%>%
  arrange(desc(total_rate_per_pop))
high_state<-state_total_rate$state[1]
low_state<-state_total_rate$state[nrow(state_total_rate)]
cat("Highest:",high_state, "\nLowest:",low_state, "\n")
```

Trajectories over time: Rhode Island fluctuates more obviously than Maine, Rhode Island also has a main spike, which is way higher than Maine.

```{r}
selected_states <- dt %>%
  filter(state %in% c(high_state, low_state))
#ggplot(selected_states, aes(x=date,y=cases_avg_per_100k,color=state)) +
  #geom_line(size = 1) 
plot_ly(data=selected_states,x=~date,y=~cases_avg_per_100k,
  color=~state,
  type='scatter',
  mode='lines',
  line=list(width = 1))
```

#### (c)

The five states which encountered the first spike is what we're looking for.

```{r}
dt20 <- dt %>%
  filter(year(date) == 2020) %>%
  select(state, date, cases_avg_per_100k)
```

```{r}
plot_ly(
  data = dt20,
  x = ~date,
  y = ~cases_avg_per_100k,
  color = ~state,
  type = "scatter",
  mode = "lines"
) %>%
  layout(
    yaxis = list(title = "COVID cases per 100k"),
    showlegend = FALSE
  )
```

The first spike may happen around March to May.

```{r}
dt20 <- dt20 %>%
  filter(month(date) %in% c(3, 4, 5)) 
plot_ly(
  data = dt20,
  x = ~date,
  y = ~cases_avg_per_100k,
  color = ~state,
  type = "scatter",
  mode = "lines"
) %>%
  layout(
    yaxis = list(title = "COVID cases per 100k")
  )
```

Take a closer look.

```{r}
dt20<-dt20 %>%
   filter(cases_avg_per_100k > 20)
plot_ly(
  data = dt20,
  x = ~date,
  y = ~cases_avg_per_100k,
  color = ~state,
  type = "scatter",
  mode = "lines"
) %>%
  layout(
    yaxis = list(title = "COVID cases per 100k")
  )
```

The first five states to experience Covid in a substantial way are New York, New Jersey, Guam, Louisana and Rhode Island.
